---
title: "Models"
author: "Group 4"
date: "`r Sys.Date()`"
output: pdf_document
---

# Load library and data
```{r message=FALSE, warning=FALSE}
setwd("/Users/atnguyen/Downloads/4.SSP")
library(readr)
library(caret)
df <- read.csv("final_combined_data.csv")
```

# Split the dataset into training and testing

```{r}
set.seed(42)
# Split the data 80 training - 20 testing
train_ind <- createDataPartition(df$livability_score, p = .8, 
                                  list = FALSE, 
                                  times = 1)
df_train <- df[train_ind, ]
df_test  <- df[-train_ind, ]
```

# Model regressing liveability on distance without considering feature of neighborhood
This model regress the liveability score on the distance of commercial spaces and does not consider characteristics such as age, income.

```{r}
# Model with distance only
lm_distances_no_feature <- lm(livability_score ~ . - Standard_Deviation_Age -Weighted_Average_Age -age_cluster -social_cluster -demographic_cluster -id - postal_code -prop_0.14 -prop_15.24 -prop_25.44 -prop_45.64 -prop_65. -prop_university  -worker - prop_women - pop_density - Aid - Income - single_family_housing, data = df_train)
summary(lm_distances_no_feature)

# Make prediction for livability_score on testing data
lm_distances_no_feature_pred <- predict(lm_distances_no_feature, newdata = df_test)

# RMSE
rmse_dis_no_fea <- sqrt(mean((lm_distances_no_feature_pred - df_test$livability_score)^2))
print(paste("RMSE:", rmse_dis_no_fea))


```


# Models considering features of neighborhood
This model regress liveability score on distance of commercial space and the feature of neighborhood. Therefore, it allows to control the characteristics of the neighborhood such as age, income. Notice that the `age` variable is recorded into 2 methods. The first method uses the average and SD value for age of citizens living in the neighborhood. The second method records age of different groups and transfer to proportion. There is 5 bins of age values `prop_0.14`, `prop_15.24`, `prop_25.44`, `prop_45.64`, and `prop_65.`.

In order to choose the best performance model, we run the model in both method recoding age.

```{r , warning=FALSE}
# Linear model with proportion of age
lm_all_age_pro <- lm(livability_score ~ . -id - postal_code - Standard_Deviation_Age -Weighted_Average_Age - age_cluster -social_cluster -demographic_cluster, data = df_train)
summary(lm_all_age_pro)

# Make prediction for livability_score on testing data
lm_all_age_pro_pred <- predict(lm_all_age_pro, newdata = df_test)
rmse_all_age_pro <- sqrt(mean((lm_all_age_pro_pred - df_test$livability_score)^2))
print(paste("RMSE:", rmse_all_age_pro))

# Linear model with average and SD of Age
lm_all_avg_sd_age <- lm(livability_score ~ . -id - postal_code - `prop_0.14` -`prop_15.24` - `prop_25.44` - `prop_45.64` - `prop_65.`- age_cluster -social_cluster -demographic_cluster, data = df_train)
summary(lm_all_avg_sd_age)

# Make prediction for livability_score on testing data
lm_all_avg_sd_age_pred <- predict(lm_all_avg_sd_age, newdata = df_test)
rmse_all_avg_sd_age <- sqrt(mean((lm_all_avg_sd_age_pred - df_test$livability_score)^2))
print(paste("RMSE:", rmse_all_avg_sd_age))
```

# Model regress liveability on different measurement of commercial space distance

## Model regress liveability on `distance`

```{r}
# Linear model with `distance`
lm_distance <-  lm(livability_score ~ `prop_0.14` +`prop_15.24` + `prop_25.44` + `prop_45.64` + `prop_65.` + prop_university + worker + prop_women + pop_density + Aid + Income + single_family_housing + Distance.GP + Distance.Hospital + Distance.Supermarket+ Distance.Department.Store + Distance.Restaurant + Distance.Day.Care + Distance.School, data = df_train)
summary(lm_distance)

# Make prediction for livability_score on testing data
lm_distance_pred <- predict(lm_distance, newdata = df_test)

# RMSE
rmse_distance <- sqrt(mean((lm_distance_pred - df_test$livability_score)^2))
print(paste("RMSE:", rmse_distance))
```
## Model regress liveability on `Within.5.km.CommercialSpace`
```{r}
# Linear model with `Whithin 5 km`
lm_within_5 <- lm(livability_score ~ `prop_0.14` +`prop_15.24` + `prop_25.44` + `prop_45.64` + `prop_65.` + prop_university + worker + prop_women + pop_density + Aid + Income + single_family_housing + Within.5.km.GP + Within.5.km.Hospital + Within.5.km.Supermarket + Within.5.km.Department.Store + Within.5.km.Restaurant + Within.5.km.Day.Care + Within.5.km.School, data = df_train)
summary(lm_within_5)

# Make prediction for livability_score on testing data
lm_within_5_pred <- predict(lm_within_5, newdata = df_test)

# RMSE
rmse_within5 <- sqrt(mean((lm_within_5_pred - df_test$livability_score)^2))
print(paste("RMSE:", rmse_within5))
```

# Model regress liveability on clusters
It is hypothesized that neighborhoods can be categorized into distinct groups based on specific characteristic. Therefore, based on the defined clusters (`demographic_cluster`, `social_cluster`, and `age_cluster`), the model regresses liveability on distance and clusters is created. Two models uses `distance` and `cluster`; and `distance`, `Within.X.Km.CommercialSpace` and `cluster` are as follows:

```{r}
# Model - `distance` and `cluster` 
lm_cluster_distance <- lm(livability_score ~ demographic_cluster + social_cluster + age_cluster + Distance.GP + Distance.Hospital + Distance.Supermarket+ Distance.Department.Store + Distance.Restaurant + Distance.Day.Care + Distance.School, data = df_train)
summary(lm_cluster_distance)

# Make prediction for livability_score on testing data
lm_cluster_distance_pred <- predict(lm_cluster_distance, newdata = df_test)

# RMSE
rmse_cluster_distances <- sqrt(mean((lm_cluster_distance_pred - df_test$livability_score)^2))
print(paste("RMSE:", rmse_cluster_distances))

# Model - `distance`, `Within.X.Km.CommercialSpace` and `cluster`
lm_cluster_distances_within <- lm(livability_score ~ . -id - postal_code - Weighted_Average_Age -Standard_Deviation_Age -prop_0.14 -prop_15.24 -prop_25.44 -prop_45.64 -prop_65. -prop_university  -worker - prop_women - pop_density - Aid - Income - single_family_housing, data = df_train)
summary(lm_cluster_distances_within)

# Make prediction for livability_score on testing data
lm_cluster_distances_within_pred <- predict(lm_cluster_distances_within, newdata = df_test)

# RMSE
rmse_cluster_dis_within <- sqrt(mean((lm_cluster_distances_within_pred - df_test$livability_score)^2))
print(paste("RMSE:", rmse_cluster_dis_within))

```
## Comparing models

To find the best performance models, we compare 7 models in term of adjusted R squares and residual mean square error (RMSE)

```{r warning=FALSE}
# Extract adjusted R-squared values from each model
adj_r_squared_lm_dis_no_fea <- summary(lm_distances_no_feature)$adj.r.squared
adj_r_squared_lm_all_age_pro <- summary(lm_all_age_pro)$adj.r.squared
adj_r_squared_lm_all_avg_sd_age <- summary(lm_all_avg_sd_age)$adj.r.squared
adj_r_squared_lm_distance <- summary(lm_distance)$adj.r.squared
adj_r_squared_lm_within_5 <- summary(lm_within_5)$adj.r.squared
adj_r_squared_lm_cluster_dis <- summary(lm_cluster_distance)$adj.r.squared
adj_r_squared_lm_cluster_dis_within <- summary(lm_cluster_distances_within)$adj.r.squared


# Create a dataframe to store the results
model_comparison <- data.frame(
 Model = c("Distance without features","Features-distances-age propotion", "Features-distances-AgeMean-AgeSD", "Featuess-Distance", "Features-Within.5.km.CommercialSpace", "Cluster-distances", "Cluster-distances-Within.X.Km.CommercialSpace"),
  Adjusted_R_squared = c(adj_r_squared_lm_dis_no_fea, adj_r_squared_lm_all_age_pro, adj_r_squared_lm_all_avg_sd_age, adj_r_squared_lm_distance, adj_r_squared_lm_within_5, adj_r_squared_lm_cluster_dis, adj_r_squared_lm_cluster_dis_within),
  RMSE = c(rmse_dis_no_fea, rmse_all_age_pro, rmse_all_avg_sd_age, rmse_distance, rmse_within5, rmse_cluster_distances, rmse_cluster_dis_within))

# Print the table
print(model_comparison)
```

# Conclusion

The two best performance models are the models that regress liveability on features, distance, and the number of commercial space within a specific radius (1km, 3km, 5km, 10km, and 20km). The different methods of age measurement do not significantly impact the Adjusted_R_square and RMSE. Therefore, these two measurement can be used to measure age of the neighborhood without affecting the prediction performance of the models.

